Projection variability: Is the family of sentences really a family?

### 2024-02-14 writing/SuB-paper/projection-operators-SuB.tex
updated analysis scripts

    Textbooks typically present the family-of-sentences diagnostic for projection without addressing by-operator variation (e.g., \citealt{chierchia_meaning_1990}). For instance, the clausal complement of \emph{discover} is diagnosed as as projective content, if its content (here: that Julian dances salsa) follows not just from \ref{ex:simple}, but also from its variants in \ref{ex:family}, where \ref{ex:simple} has been embedded under an entailment-canceling operator, such as negation \ref{ex:neg}, polar questions \ref{ex:q}, epistemic possibility modals \ref{ex:mod}, or conditional antecedents \ref{ex:cond}.

    % In formal semantics/pragmatics, the so-called `family of sentences' diagnostic is the standard way of diagnosing whether a content is projective (e.g., \citealt{chierchia_meaning_1990}). For instance, to diagnose the content of the clausal complement of \emph{discover} as projective content, a researcher would typically observe that the content that Julian dances salsa follows not just from \ref{ex:simple}, but also from the family-of-sentences variants of \ref{ex:simple} in \ref{ex:family}, where \ref{ex:simple} has been embedded under an entailment-canceling operator, such as negation \ref{ex:neg}, polar questions \ref{ex:q}, epistemic possibility modals \ref{ex:mod}, or conditional antecedents \ref{ex:cond}.

    % (e.g., \citealt{frege_uber_1892,strawson_referring_1950,kiparsky_fact_1970,karttunen_conventional_1979}).

@@

    However, prior research is suggestive of variation in the effect of different entailment-cancelling operators on projection. For instance, \citet{karttunen_observations_1971} proposed a distinction between English factive predicates (e.g., \textit{regret}) and semi-factives (e.g., \textit{discover}).
    Based on the examples in \Next, he suggested that the CC of true factives projects across all four operators, but that of semi-factives does not always project from under polar questions, modals, or conditionals.\footnote{See \citealt{beaver_have_2010} for naturally occurring examples in which the CCs of factive predicates do not project from under the four operators, thereby calling into question the distinction between factive and semi-factive predicates. For experimental research on the distinction see \citealt{djaerv-etal2016} and \citealt{djarv_cognitive_2018}. We return to this literature in section \ref{s:analysis}.}

@@

    There have been two experimental investigations of by-operator projection variation. First, the experiment reported in \citet{smith_relationship_2014} investigated projection from under negation and antecents of conditionals for various types of English projective contents. They found that the expressive content of epithets (e.g., \textit{idiot}) and the CC of \textit{know} was more projective under negation than conditionals. In contrast, the content of appositive relative clauses and the preparatory content of \textit{win} showed the opposite pattern, and the existential presupposition of clefts showed no difference. %However, their task involved asking participants how surprised they would be to learn the content under investigation after observing the utterance. Since other factors besides speaker commitment modulate discourse (un)expectedness (see e.g., \citealt{zimmermann_grammatical_2011,tonnis_german_2021}), it is unclear that they (only) measured projection.

    Second, \citet{sieker_projective_2022}  compared the projection of German factives (\textit{wissen} `know', \textit{bereuen} `regret', \textit{ent\-hüllen} `reveal') and semi-factives (\textit{bemerken} `notice', \textit{entdecken} `discover', \textit{herausfinden} `find out') between the four operators in \ref{ex:family}. The results showed that the CCs of the predicates investigated projected more from under negation than the other three operators. contrary to what \citealt{karttunen_observations_1971} suggested, a comparison of factive and semifactive predicates did not reveal that the CCs of factive predicates project more from under polar questions, modals, or antecedents of conditionals than the CCs of nonfactive predicates.

@@

    This paper reports on the results of a set of experiments that were designed to compare projection from under the four entailment-canceling operators in \ref{ex:family}. Our experiments extend the empirical scope of prior research by investigating projection for a broader set of contents, namely the contents of the complements of the 20 English clause-embedding predicates in \ref{ex:predicates}:

>>>>

    Textbooks usually present the family-of-sentences diagnostic for projection without discussing the possibility of by-operator variation (e.g., \citealt{chierchia_meaning_1990}). For instance, the clausal complement of \emph{discover} is diagnosed as as projective content, if its content (here: that Julian dances salsa) follows not just from \ref{ex:simple}, but also from the variants in \ref{ex:family}, where \ref{ex:simple} has been embedded under an entailment-canceling operator, such as negation \ref{ex:neg}, a polar question \ref{ex:q}, the epistemic possibility modal \ref{ex:mod}, or in a conditional antecedent \ref{ex:cond}.

@@

     However, prior research is suggestive that different entailment-cancelling operators may affect projection differently. For instance, \citet{karttunen_observations_1971} proposed distinguishing between English factive predicates (e.g., \textit{regret}) and semi-factives (e.g., \textit{discover}). Based on the examples in \Next, he suggested that the content of the complement (CC) of true factives consistently projects across the four operators in \Last, but that of semi-factives does not always project from under polar questions, modals, or conditionals.%
    %
    \footnote{The factive/semi-factive distinction is called into question by reports of many naturally occurring examples where the CCs of factive predicates do not project from under the four operators (see \citealt{beaver_have_2010,de_marneffe_did_2012,demarneffe-etal-sub23}). For experimental research on the distinction, see \citealt{djaerv-etal2016} and \citealt{djarv_cognitive_2018}. We return to this literature in Section \ref{s:analysis}.
    %See \citealt{beaver_have_2010} for naturally occurring examples in which the CCs of factive predicates do not project from under the four operators, thereby calling into question the distinction between factive and semi-factive predicates. \mcdm{add CB too? We have quite a lot of naturally occurring examples for that too} For experimental research on the distinction see \citealt{djaerv-etal2016} and \citealt{djarv_cognitive_2018}. We return to this literature in section \ref{s:analysis}.
    }

@@

    There have been two experimental investigations of by-operator projection variation. First, \citet{smith_relationship_2014} investigated projection from under negation and conditional antecedents for various types of English projective contents. They found that the expressive content of epithets (e.g., \textit{idiot}) and the CC of \textit{know} was more projective under negation than conditionals. In contrast, the content of appositive relative clauses and the preparatory content of \textit{win} showed the opposite pattern, and the existential presupposition of clefts showed no difference. %However, their task involved asking participants how surprised they would be to learn the content under investigation after observing the utterance. Since other factors besides speaker commitment modulate discourse (un)expectedness (see e.g., \citealt{zimmermann_grammatical_2011,tonnis_german_2021}), it is unclear that they (only) measured projection.

    Second, \citet{sieker_projective_2022}  compared the projection of German factives (\textit{wissen} `know', \textit{bereuen} `regret', \textit{ent\-hüllen} `reveal') and semi-factives (\textit{bemerken} `notice', \textit{entdecken} `discover', \textit{herausfinden} `find out') between the four operators in \ref{ex:family}. The results showed that the CCs of the predicates investigated projected more from under negation than the other three operators. Contrary to what \citet{karttunen_observations_1971} suggested, a comparison of the factive and
    % \mcdm{need to make a pass to make sure semi-factive is written consistantly, above it is with hyphen}
    semi-factive predicates did not reveal that the CCs of factive predicates project more from under polar questions, modals, or antecedents of conditionals than the CCs of semi-factive predicates.

@@

    This paper presents the results of a set of experiments designed to compare projection from under the four entailment-canceling operators in \ref{ex:family} in English. Our experiments extend the empirical scope of prior research on by-operator projection variation by investigating a larger set of contents, namely the contents of the complements of the 20 English clause-embedding predicates in \ref{ex:predicates}:

    Our research expands on previous studies by exploring a broader range of content, specifically the complements of the 20 English clause-embedding predicates listed in (4).
